Prolfqua - Proteomics label free quantification using linear models.

### 2022-01-25 inst/MyArticle/MergePoster.Rmd
package names all lowercase, and remove library statements.

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
library(ggplot2)

@@

```{r exampleContrasts, fig.cap="(ref:exampleContrasts)", echo=TRUE, out.width = '90%'}
contrasts <- c(
  "dilution_(9/7.5)_1.2" =   "dilution.e - dilution.d",
  "dilution_(7.5/6)_1.25" =   "dilution.d - dilution.c"
  #,"dilution_(6/4.5)_1.3(3)" =   "dilution.c - dilution.b",
  #,"dilution_(4.5/3)_1.5" =   "dilution.b - dilution.a"
)
# fit model
lmmodel <- paste(lfqp$intensity_column()," ~ dilution.")
modelFunction <- strategy_lm( lmmodel, model_name = "lm")
models <- build_model(lfqp, modelFunction)
p0 <- models$anova_histogram()
# compute contrasts from linear model and with imputation
contr <- prolfqua::Contrasts$new(models, contrasts) %>% 
  prolfqua::ContrastsModerated$new()
conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts) %>% 
  prolfqua::ContrastsModerated$new()
tmp <- prolfqua::addContrastResults(prefer = contr, add = conI)
# visualize results
pl <- tmp$merged$get_Plotter()
p1 <- pl$histogram()$p.value
p2 <- pl$volcano()$FDR +
   theme(legend.position = "bottom")
gridExtra::grid.arrange(p1, p2, ncol = 2)

@@

ttt <- sapply(allBenchmarks, function(x){x$complete(FALSE)})
res <- map_df(allBenchmarks, function(x){x$pAUC()})
res <- res %>% mutate(whatfix = case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))

norm <- res %>% group_by(contrast,whatfix) %>% summarize(meanpAUC_10 = mean(pAUC_10))
res <- inner_join(res, norm)
res <- mutate(res , pAUC_10n = pAUC_10 - meanpAUC_10)

resAllB <- res %>% dplyr::filter(contrast == "all")

@@

dd <- map_df(allBenchmarks, function(x){res <- x$smc$summary; res$name <- x$model_name;res})
dd
dd <- dd %>% mutate(nrcontrasts = protein_Id * (4 - nr_missing))
dd
dds <- dd %>% group_by(name) %>% summarize(nrcontrasts = sum(nrcontrasts))
dds
dds$percent <- dds$nrcontrasts/max(dds$nrcontrasts) * 100

nrgg <- dds %>% ggplot(aes(x = name, y = nrcontrasts )) + 
  geom_bar(stat = "identity", fill="white", colour = "black") + 
  coord_cartesian(ylim = c(min(dds$nrcontrasts) - 100, max(dds$nrcontrasts) + 10)) +
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5)) +
  geom_text(aes(label = round(nrcontrasts, digits = 1)),
            vjust = 1, hjust = -0.2, angle = -90) + 
  geom_text(aes(label = paste0("(",round(percent, digits = 1),"%)")),
            vjust = -1, hjust = -0.2, angle = -90) 
p3 <- nrgg

@@

ddb <- filter(dd, contrast == "dilution_(4.5/3)_1.5")
ddb <- dd %>% dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
ddb <- dd %>% dplyr::filter(contrast == "all")

p4 <- ddb %>% ggplot(aes(y = FDP_,  x  = scorecol )) + 
  geom_line(aes(color = model_name, linetype = model_name)) +
  facet_wrap(~contrast) + 
   geom_abline(intercept = 0, slope = 1, color = 2) + 
   theme(legend.position = "bottom")

@@

```{r echo=TRUE, eval=FALSE}
install.packages('remotes')
remotes::install_gitlab("wolski/prolfquaData",
                        host="gitlab.bfabric.org")
remotes::install_github('wolski/prolfqua',
                        build_vignettes = TRUE)

@@

tmp <- inner_join(prodSum$ftable$content, modSum$ftable$content, by=c("contrast","what"), suffix = c(".proDA",".lmMod"))
tmp <- tmp %>% unite("score_contr" , what, contrast, remove = FALSE)

>>>>

```{r setup, include=FALSE}

@@

```{r exampleContrasts, fig.cap="(ref:exampleContrasts)", echo=TRUE, out.width = '90%'}
contrasts <- c(
  "dilution_(9/7.5)_1.2" =   "dilution.e - dilution.d",
  "dilution_(7.5/6)_1.25" =   "dilution.d - dilution.c"
  #,"dilution_(6/4.5)_1.3(3)" =   "dilution.c - dilution.b",
  #,"dilution_(4.5/3)_1.5" =   "dilution.b - dilution.a"
)
# fit model
lmmodel <- paste(lfqp$intensity_column()," ~ dilution.")
modelFunction <- strategy_lm( lmmodel, model_name = "lm")
models <- build_model(lfqp, modelFunction)
p0 <- models$anova_histogram()
# compute contrasts from linear model and with imputation
contr <- prolfqua::Contrasts$new(models, contrasts) |> 
  prolfqua::ContrastsModerated$new()
conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts) |> 
  prolfqua::ContrastsModerated$new()
tmp <- prolfqua::addContrastResults(prefer = contr, add = conI)
# visualize results
pl <- tmp$merged$get_Plotter()
p1 <- pl$histogram()$p.value
p2 <- pl$volcano()$FDR +
   theme(legend.position = "bottom")
gridExtra::grid.arrange(p1, p2, ncol = 2)

@@

ttt <- sapply(allBenchmarks, function(x){x$complete(FALSE)})
res <- map_df(allBenchmarks, function(x){x$pAUC()})
res <- res |> mutate(whatfix = case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))

norm <- res |> group_by(contrast,whatfix) |> summarize(meanpAUC_10 = mean(pAUC_10))
res <- inner_join(res, norm)
res <- mutate(res , pAUC_10n = pAUC_10 - meanpAUC_10)

resAllB <- res |> dplyr::filter(contrast == "all")

@@

dd <- map_df(allBenchmarks, function(x){res <- x$smc$summary; res$name <- x$model_name;res})
dd
dd <- dd |> mutate(nrcontrasts = protein_Id * (4 - nr_missing))
dd
dds <- dd |> group_by(name) |> summarize(nrcontrasts = sum(nrcontrasts))
dds
dds$percent <- dds$nrcontrasts/max(dds$nrcontrasts) * 100

nrgg <- dds |> ggplot(aes(x = name, y = nrcontrasts )) + 
  geom_bar(stat = "identity", fill="white", colour = "black") + 
  coord_cartesian(ylim = c(min(dds$nrcontrasts) - 100, max(dds$nrcontrasts) + 10)) +
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5)) +
  geom_text(aes(label = round(nrcontrasts, digits = 1)),
            vjust = 1, hjust = -0.2, angle = -90) + 
  geom_text(aes(label = paste0("(",round(percent, digits = 1),"%)")),
            vjust = -1, hjust = -0.2, angle = -90) 
p3 <- nrgg

@@

ddb <- filter(dd, contrast == "dilution_(4.5/3)_1.5")
ddb <- dd |> dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
ddb <- dd |> dplyr::filter(contrast == "all")

p4 <- ddb |> ggplot(aes(y = FDP_,  x  = scorecol )) + 
  geom_line(aes(color = model_name, linetype = model_name)) +
  facet_wrap(~contrast) + 
   geom_abline(intercept = 0, slope = 1, color = 2) + 
   theme(legend.position = "bottom")

@@

```{r echo=TRUE, eval=FALSE}
install.packages('remotes')
remotes::install_gitlab("wolski/prolfquadata",
                        host="gitlab.bfabric.org")
remotes::install_github('wolski/prolfqua',
                        build_vignettes = TRUE)

@@

tmp <- inner_join(prodSum$ftable$content, modSum$ftable$content, by=c("contrast","what"), suffix = c(".proDA",".lmMod"))
tmp <- tmp |> unite("score_contr" , what, contrast, remove = FALSE)

### 2022-01-25 inst/MyArticle/PosterBioconductor.Rmd
package names all lowercase, and remove library statements.

```{r setup, include=FALSE}
library(tidyverse)

library(ggplot2)

@@

```{r exampleContrasts, fig.cap="(ref:exampleContrasts)", echo=TRUE, out.width = '90%'}
contrasts <- c(
  "dilution_(9/7.5)_1.2" =   "dilution.e - dilution.d",
  "dilution_(7.5/6)_1.25" =   "dilution.d - dilution.c",
  "dilution_(6/4.5)_1.3(3)" =   "dilution.c - dilution.b",
  "dilution_(4.5/3)_1.5" =   "dilution.b - dilution.a"
)
# fit model
lmmodel <- paste(lfqp$intensity_column()," ~ dilution.")
modelFunction <- strategy_lm( lmmodel, model_name = "lm")
models <- build_model(lfqp, modelFunction)
p1 <-  models$anova_histogram()
# compute contrasts from linear model and with imputation
contr <- Contrasts$new(models, contrasts) %>% 
  ContrastsModerated$new()
conI <- ContrastsSimpleImpute$new( lfqp, contrasts) %>% 
  ContrastsModerated$new()
tmp <- addContrastResults(contr, conI)
# visualize results
pl <- tmp$merged$get_Plotter()
p1 <- pl$histogram()$p.value
p2 <- pl$volcano()$FDR +
   theme(legend.position = "bottom")
gridExtra::grid.arrange(p1, p2, ncol = 2)

@@

dd <- map_df(allBenchmarks, function(x){res <- x$smc$summary; res$name <- x$model_name;res})
dd <- dd %>% mutate(nrcontrasts = protein_Id * (4 - nr_missing))
dds <- dd %>% group_by(name) %>% summarize(nrcontrasts = sum(nrcontrasts))
dds$percent <- dds$nrcontrasts/max(dds$nrcontrasts) * 100

nrgg <- dds %>% ggplot(aes(x = name, y = nrcontrasts )) + 
  geom_bar(stat = "identity", fill="white", colour = "black") + 
  coord_cartesian(ylim = c(min(dds$nrcontrasts) - 100, max(dds$nrcontrasts) + 10)) +
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5)) +
  geom_text(aes(label = round(nrcontrasts, digits = 1)),
            vjust = 1, hjust = -0.2, angle = -90) + 
  geom_text(aes(label = paste0("(",round(percent, digits = 1),"%)")),
            vjust = -1, hjust = -0.2, angle = -90) 
p1 <- nrgg

@@

ddb <- filter(dd, contrast == "dilution_(4.5/3)_1.5")
ddb <- dd %>% dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
ddb <- dd %>% dplyr::filter(contrast == "all")

p2 <- ddb %>% ggplot(aes(y = FDP_,  x  = scorecol )) + 
  geom_line(aes(color = model_name, linetype = model_name)) +
  facet_wrap(~contrast) + 
   geom_abline(intercept = 0, slope = 1, color = 2) + 
   theme(legend.position = "bottom")

@@

```{r echo=TRUE, eval=FALSE}
install.packages('remotes')
remotes::install_gitlab("wolski/prolfquaData",
                        host="gitlab.bfabric.org")
remotes::install_github('wolski/prolfqua',
                        build_vignettes = TRUE)
```

>>>>

```{r setup, include=FALSE}

@@

```{r exampleContrasts, fig.cap="(ref:exampleContrasts)", echo=TRUE, out.width = '90%'}
contrasts <- c(
  "dilution_(9/7.5)_1.2" =   "dilution.e - dilution.d",
  "dilution_(7.5/6)_1.25" =   "dilution.d - dilution.c",
  "dilution_(6/4.5)_1.3(3)" =   "dilution.c - dilution.b",
  "dilution_(4.5/3)_1.5" =   "dilution.b - dilution.a"
)
# fit model
lmmodel <- paste(lfqp$intensity_column()," ~ dilution.")
modelFunction <- strategy_lm( lmmodel, model_name = "lm")
models <- build_model(lfqp, modelFunction)
p1 <-  models$anova_histogram()
# compute contrasts from linear model and with imputation
contr <- Contrasts$new(models, contrasts) |> 
  ContrastsModerated$new()
conI <- ContrastsSimpleImpute$new( lfqp, contrasts) |> 
  ContrastsModerated$new()
tmp <- addContrastResults(contr, conI)
# visualize results
pl <- tmp$merged$get_Plotter()
p1 <- pl$histogram()$p.value
p2 <- pl$volcano()$FDR +
   theme(legend.position = "bottom")
gridExtra::grid.arrange(p1, p2, ncol = 2)

@@

dd <- map_df(allBenchmarks, function(x){res <- x$smc$summary; res$name <- x$model_name;res})
dd <- dd |> mutate(nrcontrasts = protein_Id * (4 - nr_missing))
dds <- dd |> group_by(name) |> summarize(nrcontrasts = sum(nrcontrasts))
dds$percent <- dds$nrcontrasts/max(dds$nrcontrasts) * 100

nrgg <- dds |> ggplot(aes(x = name, y = nrcontrasts )) + 
  geom_bar(stat = "identity", fill="white", colour = "black") + 
  coord_cartesian(ylim = c(min(dds$nrcontrasts) - 100, max(dds$nrcontrasts) + 10)) +
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5)) +
  geom_text(aes(label = round(nrcontrasts, digits = 1)),
            vjust = 1, hjust = -0.2, angle = -90) + 
  geom_text(aes(label = paste0("(",round(percent, digits = 1),"%)")),
            vjust = -1, hjust = -0.2, angle = -90) 
p1 <- nrgg

@@

ddb <- dplyr::filter(dd, contrast == "dilution_(4.5/3)_1.5")
ddb <- dd |> dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
ddb <- dd |> dplyr::filter(contrast == "all")

p2 <- ddb |> ggplot(aes(y = FDP_,  x  = scorecol )) + 
  geom_line(aes(color = model_name, linetype = model_name)) +
  facet_wrap(~contrast) + 
   geom_abline(intercept = 0, slope = 1, color = 2) + 
   theme(legend.position = "bottom")

@@

```{r echo=TRUE, eval=FALSE}
install.packages('remotes')
remotes::install_gitlab("wolski/prolfquadata",
                        host="gitlab.bfabric.org")
remotes::install_github('wolski/prolfqua',
                        build_vignettes = TRUE)
```

### 2022-01-25 inst/MyArticle/Prolfqua.Rmd
package names all lowercase, and remove library statements.

```{r setup, include=FALSE}
library(tidyverse)
library(prolfqua)
library(ggplot2)

@@

tt <- data.frame(contrastsName = names(Contrasts),  contrasts = Contrasts)
tt <- tt %>% separate(contrastsName, c("dilution", "ratio", "expected fold change"), sep =  "_", remove = FALSE)

@@

tt <- tt %>% dplyr::filter(contrastsName  %in% relevantContrasts)
tt <- select(tt, all_of(c("contrasts","ratio", "expected fold change")))
knitr::kable(tt, caption = "Contrasts used for benchmark.", format = "latex")

>>>>

```{r setup, include=FALSE}

@@

tt <- data.frame(contrastsName = names(Contrasts),  contrasts = Contrasts)
tt <- tt |> separate(contrastsName, c("dilution", "ratio", "expected fold change"), sep =  "_", remove = FALSE)

@@

tt <- tt |> dplyr::filter(contrastsName  %in% relevantContrasts)
tt <- select(tt, all_of(c("contrasts","ratio", "expected fold change")))
knitr::kable(tt, caption = "Contrasts used for benchmark.", format = "latex")

### 2022-01-25 inst/MyArticle/ProlfquaPlain.Rmd
package names all lowercase, and remove library statements.

tt <- tt %>% dplyr::filter(contrastsName  %in% relevantContrasts)
tt <- dplyr::select(tt, all_of(c("contrasts","ratio", "expected fold change")))
knitr::kable(tt, caption = "Contrasts used for benchmark.")

@@

dd <- map_df(allBenchmarks, function(x){res <- x$smc$summary; res$name <- x$model_name;res})
dd <- dd %>% mutate(nrcontrasts = protein_Id * (4 - nr_missing))
dds <- dd %>% group_by(name) %>% summarize(nrcontrasts = sum(nrcontrasts))
dds$percent <- dds$nrcontrasts/max(dds$nrcontrasts) * 100

nrgg <- dds %>% ggplot(aes(x = name, y = nrcontrasts )) + 
  geom_bar(stat = "identity", fill="white", colour = "black") + 
  coord_cartesian(ylim = c(min(dds$nrcontrasts) - 100, max(dds$nrcontrasts) + 10)) +
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5)) +
  geom_text(aes(label = round(nrcontrasts, digits = 1)),
            vjust = 1, hjust = -0.2, angle = -90) + 
  geom_text(aes(label = paste0("(",round(percent, digits = 1),"%)")),
            vjust = -1, hjust = -0.2, angle = -90) 
nrgg

@@

ttt <- sapply(allBenchmarks, function(x){x$complete(FALSE)})
res <- map_df(allBenchmarks, function(x){x$pAUC()})
res <- res %>% mutate(whatfix = case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))

res$contrast %>% unique()
#resAllB <- res %>% dplyr::filter(contrast == "dilution_(9/7.5)_1.2")
#resAllB <- res %>% dplyr::filter(contrast == "dilution_(4.5/3)_1.5")
#resAllB <- res %>% dplyr::filter(contrast == "dilution_(7.5/6)_1.25")

norm <- res %>% group_by(contrast,whatfix) %>% summarize(meanpAUC_10 = mean(pAUC_10))
res <- inner_join(res, norm)
res <- mutate(res , pAUC_10n = pAUC_10 - meanpAUC_10)

resAllB <- res %>% dplyr::filter(contrast == "all")

@@

```{r FDRfdp, fig.cap = "Compare FDR estimate with false discovery proportion (FDP).", fig.width=8, fig.height=8}
dd <- map_df(allBenchmarks, function(x){res <- x$get_confusion_FDRvsFDP(); res$name <- x$model_name;res})
ddb <- filter(dd, contrast == "dilution_(4.5/3)_1.5")
ddb <- dd %>% dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
ddb <- dd %>% dplyr::filter(contrast == "all")

ddb %>% ggplot(aes(y = FDP_,  x  = scorecol )) + 
  geom_line(aes(color = model_name, linetype = model_name)) +
  facet_wrap(~contrast) + 
   geom_abline(intercept = 0, slope = 1, color = 2)

@@

```{r fig.cap="Distribution of the FDR values produced."}
dd %>% ggplot(aes(x = scorecol)) + geom_histogram() + facet_wrap(~model_name)
```

>>>>

tt <- tt |> dplyr::filter(contrastsName  %in% relevantContrasts)
tt <- dplyr::select(tt, all_of(c("contrasts","ratio", "expected fold change")))
knitr::kable(tt, caption = "Contrasts used for benchmark.")

@@

dd <- map_df(allBenchmarks, function(x){res <- x$smc$summary; res$name <- x$model_name;res})
dd <- dd |> mutate(nrcontrasts = protein_Id * (4 - nr_missing))
dds <- dd |> group_by(name) |> summarize(nrcontrasts = sum(nrcontrasts))
dds$percent <- dds$nrcontrasts/max(dds$nrcontrasts) * 100

nrgg <- dds |> ggplot(aes(x = name, y = nrcontrasts )) + 
  geom_bar(stat = "identity", fill="white", colour = "black") + 
  coord_cartesian(ylim = c(min(dds$nrcontrasts) - 100, max(dds$nrcontrasts) + 10)) +
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5)) +
  geom_text(aes(label = round(nrcontrasts, digits = 1)),
            vjust = 1, hjust = -0.2, angle = -90) + 
  geom_text(aes(label = paste0("(",round(percent, digits = 1),"%)")),
            vjust = -1, hjust = -0.2, angle = -90) 
nrgg

@@

ttt <- sapply(allBenchmarks, function(x){x$complete(FALSE)})
res <- map_df(allBenchmarks, function(x){x$pAUC()})
res <- res |> mutate(whatfix = case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))

res$contrast |> unique()
#resAllB <- res |> dplyr::filter(contrast == "dilution_(9/7.5)_1.2")
#resAllB <- res |> dplyr::filter(contrast == "dilution_(4.5/3)_1.5")
#resAllB <- res |> dplyr::filter(contrast == "dilution_(7.5/6)_1.25")

norm <- res |> group_by(contrast,whatfix) |> summarize(meanpAUC_10 = mean(pAUC_10))
res <- inner_join(res, norm)
res <- mutate(res , pAUC_10n = pAUC_10 - meanpAUC_10)

resAllB <- res |> dplyr::filter(contrast == "all")

@@

```{r FDRfdp, fig.cap = "Compare FDR estimate with false discovery proportion (FDP).", fig.width=8, fig.height=8}
dd <- map_df(allBenchmarks, function(x){res <- x$get_confusion_FDRvsFDP(); res$name <- x$model_name;res})
ddb <- filter(dd, contrast == "dilution_(4.5/3)_1.5")
ddb <- dd |> dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
ddb <- dd |> dplyr::filter(contrast == "all")

ddb |> ggplot(aes(y = FDP_,  x  = scorecol )) + 
  geom_line(aes(color = model_name, linetype = model_name)) +
  facet_wrap(~contrast) + 
   geom_abline(intercept = 0, slope = 1, color = 2)

@@

```{r fig.cap="Distribution of the FDR values produced."}
dd |> ggplot(aes(x = scorecol)) + geom_histogram() + facet_wrap(~model_name)
```
